\documentclass[12pt]{ucthesis}
\newif \ifpdf
\ifx \pdfoutput \undefined
  \pdffalse
\else
  \pdfoutput=1
  \pdftrue
\fi
\usepackage{url}
\ifpdf
  \usepackage[pdftex]{graphicx}
  \usepackage[pdftex,
    plainpages=false,
    breaklinks=true,
    colorlinks=true,
    urlcolor=blue,
    citecolor=blue,
    linkcolor=blue,
    bookmarks=true,
    bookmarksopen=true,
    bookmarksopenlevel=3,
    pdfstartview=FitV,
    pdfauthor={William Faught},
    pdftitle={Interoperation for Incompatible Evaluation Strategies},
    pdfkeywords={thesis, masters, cal poly}]{hyperref}
  \pdfcompresslevel=1
\else
  \usepackage{graphicx}
\fi
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[letterpaper]{geometry}
\usepackage{mathrsfs}
\usepackage{setspace}
\usepackage[overload]{textcase}
\usepackage{array}
\usepackage{xspace}
\usepackage{color}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{case}{Case}[theorem]
\newtheorem{subcase}{Case}[case]
\bibliographystyle{abbrv}
\setlength{\parindent}{0.25in}
\setlength{\parskip}{6pt}
\setlength{\extrarowheight}{5pt}
\geometry{verbose,nohead,tmargin=1.25in,bmargin=1in,lmargin=1.5in,rmargin=1.3in}
\setcounter{tocdepth}{2}
\newcommand{\captionfonts}{\small\bf\ssp}
\makeatletter
\long\def\@makecaption#1#2{%
  \vskip\abovecaptionskip
  \sbox\@tempboxa{{\captionfonts #1: #2}}
  \ifdim \wd \@tempboxa > \hsize
    {\captionfonts #1: #2\par}
  \else
    \hbox to \hsize{\hfil\box\@tempboxa\hfil}
  \fi
  \vskip\belowcaptionskip}
%\let\@currsize\normalsize
\makeatother
\begin{document}
\title{Interoperation for Incompatible Evaluation Strategies}
\author{William Faught}
\degreemonth{June}
\degreeyear{2008}
\degree{Master of Science}
\defensemonth{August}
\defenseyear{2008}
\numberofmembers{3}
\chair{Dr. John Clements}
\othermemberA{Dr. Gene Fisher}
\othermemberB{Dr. Aaron Keen}
\field{Computer Science}
\campus{San Luis Obispo}
\copyrightyears{seven}
\maketitle
\begin{frontmatter}
\copyrightpage
\approvalpage
\begin{abstract}
Software components written in different programming languages can cooperate through interoperation.  Differences between languages---incompatibilities---complicate interoperation.  This paper explores and resolves incompatible type systems, support for parametricity, and evaluation strategies with a model of computation, gives a thorough proof of its type soundness, and describes an implementation of it.  The model uses contracts for higher-order functions and lump types to resolve incompatible type systems, label types to resolve incompatible support for parametricity, and delayed conversions for list constructions to resolve incompatible evaluation strategies.  These mechanisms enable the interoperation of Haskell, ML, and Scheme without compromising their semantics.
\end{abstract}
\begin{acknowledgements}
\indent\indent I want to thank my parents, Jerry and Jo Ann, for their encouragement, advice, and support, without which this would not have been possible.

I want to thank my adviser, John Clements, for helping me along the way.  I very much appreciate the time he set aside for me and his advice.
\end{acknowledgements}
\tableofcontents
\listoffigures
\end{frontmatter}
\pagestyle{plain}
\renewcommand{\baselinestretch}{1.66}
\input{macros.tex}
\input{chapter-1.tex}
\input{chapter-2.tex}
\input{chapter-3.tex}
\input{chapter-4.tex}
\input{chapter-5.tex}
\input{chapter-6.tex}
\clearpage
\bibliography{bibliography}
\bibliographystyle{plain}
\end{document}